---
title: "52414 - lab 1 Solutions"
author: "Halit Cohen and Yoray Mauda"
date: "25/4/2022"
output: html_document
---


```{r, include=FALSE}
library(tidyverse) # This includes dplyr, stringr, ggplot2, ..
library(dplyr)
library(data.table)
library(rworldmap) # world map
library(ggthemes)
library(reshape2) # melt: change data-frame format long/wide
library(e1071) # skewness and kurtosis
```  

**Solution:**  

1. In this section, we will load the data frame, and will compute basic processing. Plus, we will list a table the top five dates as requested. 
```{r} 

# a. Loading the data:
df <- read.csv("C:\\Users\\halitcohen\\Desktop\\second year\\R\\owid-covid-data.csv")

# Modifying date because its in "character" class.
df$date <- as.Date(df$date,origin = "%Y-%m-%d")
class(df$date)
# b. Listing five top dates
High_income_new_cases <- filter(df,location == "High income")%>%
arrange(desc(new_cases))
top_n(High_income_new_cases[c("date","new_cases")],5)

High_income_new_deaths <- filter(df,location == "High income")%>%
  arrange(desc(new_deaths))
top_n(High_income_new_cases[c("date","new_deaths")],5)

High_income_new_vaccinations <- filter(df,location == "High income")%>%
  arrange(desc(new_vaccinations))
top_n(High_income_new_cases[c("date","new_vaccinations")],5)

```
[Solution text - description of analysis and results]


2. In this Question, we will compute a function that plots high and low income in one graph, as a function of date. 
```{r}
f <- function(data_frame, col_name) {
filtered_covid_data_high <- data_frame %>% filter(location == "High income")
filtered_covid_data_low <- data_frame %>% filter(location == "Low income")
plot(filtered_covid_data_high$date, filtered_covid_data_high$col_name , xlab = "Date", ylab = "amount", col = "red")
points(filtered_covid_data_low$date, filtered_covid_data_low$col_name, col = "blue")
}
f(df,new_deaths)
```
[Solution text - description of analysis and results]

3. [Solution text - explanations] 
```{r} 
# Solution code in R 

```
[Solution text - description of analysis and results]

4. [Solution text - explanations] 
```{r} 
# Solution code in R 

```
[Solution text - description of analysis and results]

5. [Solution text - explanations] 
```{r} 
# Solution code in R 

```
[Solution text - description of analysis and results]

6. [Solution text - explanations] 
```{r} 
# Solution code in R 

```
[Solution text - description of analysis and results]

7. [Solution text - explanations] 
```{r} 
# Solution code in R 

```
[Solution text - description of analysis and results]

8. [Solution text - explanations] 
```{r} 
# Solution code in R 

```
[Solution text - description of analysis and results]

9. [Solution text - explanations] 
```{r} 
# Solution code in R 

```
[Solution text - description of analysis and results]

10. [Solution text - explanations] 
```{r} 
# Solution code in R 

```
[Solution text - description of analysis and results]

11. [Solution text - explanations] 
```{r} 
# Solution code in R 

```
[Solution text - description of analysis and results]

<br/><br/>  



